The \dealii Library, Version 9.5

### 2023-06-21 9.5/paper.tex
Add PETSc stuff

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}\xspace}
\newcommand{\pfrst}{{\specialword{p4est}}\xspace}
\newcommand{\trilinos}{{\specialword{Trilinos}}\xspace}
\newcommand{\aspect}{\specialword{Aspect}\xspace}
\newcommand{\petsc}{\specialword{PETSc}\xspace}
\newcommand{\cmake}{{\specialword{CMake}}\xspace}
\newcommand{\candi}{{\specialword{candi}}\xspace}

@@

\todo[inline]{Stefano Z.: Please write this section.}

\begin{itemize}
\item PreconditionBDDC
\item SNES
\item TS
\item Wrapping PETSc matrices and vectors
\end{itemize}

>>>>

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}\xspace}
\newcommand{\pfrst}{{\specialword{p4est}}\xspace}
\newcommand{\trilinos}{{\specialword{Trilinos}}\xspace}
\newcommand{\aspect}{\specialword{Aspect}\xspace}
\newcommand{\petsc}{\specialword{PETSc}\xspace}
\newcommand{\snes}{{\specialword{SNES}}\xspace}
\newcommand{\ts}{{\specialword{TS}}\xspace}
\newcommand{\petscsf}{{\specialword{SF}}\xspace}
\newcommand{\cmake}{{\specialword{CMake}}\xspace}
\newcommand{\candi}{{\specialword{candi}}\xspace}
\newcommand{\sundials}{{\specialword{SUNDIALS}}\xspace}
\newcommand{\kinsol}{{\specialword{KINSOL}}\xspace}
\newcommand{\ida}{{\specialword{IDA}}\xspace}
\newcommand{\arkode}{{\specialword{ARKODE}}\xspace}

@@

The \dealii classes wrapping \petsc objects have been heavily rewritten
to support the nonlinear solver \snes and the Ordinary Differential Equations (ODE)
solver \ts \cite{abhyankar2018petsc}.
First, we briefly describe the most important improvements to the existing
classes and then outline the newly designed interfaces to the nonlinear solvers, together
with a new interface to the communication module \petscsf in \petsc \cite{zhang2021petscsf}.

Vector and matrix classes of the \petsc wrappers have been extended with an additional constructor
that can wrap an already existing \petsc vector or matrix, respectively.
The \texttt{BlockVector} and
\texttt{BlockSparseMatrix} classes now internally use
\petsc nested objects, i.e. \texttt{VECNEST} and  \texttt{MATNEST} respectively.
We have added a new class \texttt{PETScWrappers::PreconditionShell} to support user-defined
preconditioning that can be simply customized as:
\begin{c++}
PETScWrappers::PreconditionShell preconditioner(/*...*/);

preconditioner.vmult = [&](const VectorType &src,
                           VectorType       &dst) {/*...*/};

\end{c++}
The resulting object can be passed to \petsc and used within the nonlinear solver hierarchy.
See \ref{sec:callbacks} for additional informations on such kind of callbacks.

The \petsc~\snes class solves system of nonlinear equations of the form $F(x) = 0$.
The interface to \snes has been modeled on the already existing interface
to the \kinsol solver from the \sundials package, and the nonlinear problem
is specified with a set of callbacks that are used by the solver:
\begin{c++}
PETScWrappers::NonlinearSolver<VectorType> solver(/*...*/);

solver.residual            = [&](const VectorType &x,
                                 VectorType       &F) {/*...*/};
solver.setup_jacobian      = [&](const VectorType &x) {/*...*/};
solver.solve_with_jacobian = [&](const VectorType &rhs,
                                 VectorType       &sol) {/*...*/};
solver.solve(x);
\end{c++}
The default configuration is set up to use a Jacobian-free Newton-Krylov (JFNK)
approach~\cite{knoll2004jacobian}, using \texttt{solve\_with\_jacobian} as linear solver.
Numerous other solver configurations are possible and can be selected
programmatically, or via the powerful command line customization of \petsc.
This includes for example Quasi-Newton methods, Anderson's acceleration, and
nonlinear preconditioning~\cite{brune2015composing}.

The \petsc~\ts class solves ODEs in explicit or implicit form \cite{abhyankar2018petsc}, i.e.:
\begin{eqnarray*}
%\frac{\partial u}{\partial t} = G(t,u), &\text{explicit}\\
%F(t,u,\frac{\partial u}{\partial t}) = 0, &\text{implicit}\\
\dot{u} = G(t,u), &\text{(explicit)}\\
F(t,u,\dot{u}) = 0, &\text{(implicit)}.\\
\end{eqnarray*}
The interface to \ts has been modeled on the already existing interfaces
to the \ida and \arkode solvers from the \sundials package. Specifically:
\begin{c++}
PETScWrappers::TimeDependentSolver<VectorType> solver(/*...*/);

// If solving udot = G(t,u)
solver.explicit_function = [&](const double     t,
                               const VectorType &u,
                               VectorType       &G) {/*...*/};
// If solving F(t,u,udot) = 0
solver.implicit_function = [&](const double     t,
                               const VectorType &u,
                               const VectorType &udot,
                               VectorType       &F) {/*...*/};
\end{c++}
In addition to specifying the function callbacks, users can further customize
the solution of the linearized equations $\alpha * \frac{\partial{F}}{\partial{\dot{u}}} + \frac{\partial{F}}{\partial{u}}$
via the callbacks:
\begin{c++}
solver.setup_jacobian = [&](const double     t,
                            const VectorType &u,
                            const VectorType &udot,
                            const double     alpha) {/*...*/};
solver.solve_with_jacobian = [&](const VectorType &rhs,
                                 VectorType       &sol) {/*...*/};
\end{c++}
As for \snes, the default configuration of an implicit solver is set up to
a JFNK approach, and the entire suite of solvers offered by PETSc is available,
including adaptive time-stepping and Implicit-Explicit schemes.

We refer interested readers to our documentation for other more advanced
functions of the \ts and \snes wrappers, and to the {\tt tests/petsc/} folder
for examples on how to use them.

We close this section by introducing the interface to the \petscsf class,
the abstract communication model of \petsc. The \dealii interface to
\petscsf has been modelled on those of \texttt{Utilities::MPI::Partitioner} and
\texttt{Utilities::MPI::NoncontiguousPartitioner}. Future developments
will add support for GPU buffers.
